---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# CDOMtools

CDOMtools provides various functions for processing UV-visible absorption spectra for analysis of chromophoric dissolved organic matter (CDOM).

## Installation

You can install the development version of CDOMtools from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
devtools::install_github("bristol-emily/CDOMtools")

## load library
library(CDOMtools)

```

## Example: Calculate Napierian absorption coefficients

An example for converting optical density/absorbance to Napierian absorption coefficients ($m^{-1}$).

```{r convert units}

## load example absorption spectra dataset
data("example")

## create a new column that contains napierian absorption coefficients
example$absorption <- calc_absorption_coef(example$absorption, # vector of absorbance values
                                   method = "napierian", # specify napierian or decadic units
                                   path_length = 0.01) # path length (i.e. cuvette diameter)
  
```

## Example: Calculate spectral slope

Calculate the log-transformed spectral slope in the 275-295 nm wavelength region.

```{r spectral slope}

slope <- calc_spectral_slope(example$wavelength, # vector of wavelength values (nm)
                             example$absorption, # vector of absorption values (m^-1)
                             start = 275, # start of wavelength region
                             end = 295) # end of wavelength region

```

## Example: Calculate specific ultraviolet absorbance

Calculate specific ultraviolet absorbance (decadic units) normalized by DOC concentration in $mg L^-1$ (e.g. $SUVA_254$).

```{r SUVA}

SUVA254 <- calc_SUVA(example$wavelength, # vector of wavelength values (nm)
                     example$absorption, # vector of absorption values 
                     11.3, # DOC concentration
                     specific_wavelength = 254, # specific wavelength of interest 
                     absorption_coef = "napierien", # unit of absorption value provided
                     DOC_unit = "mg") # unit of DOC value provided

```

## Vignettes

See the following vignette for an example workflow to process batches of CDOM data aquired using OceanView software:

``` r
browseVignettes("oceanview-workflow")
```
