---
title: "oceanview-workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{oceanview-workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

## Processing raw data from OceanOptics OceanView software

The code in this vignette is written to process UV-vis absorption spectra taken with an OceanOptics mini spec using OceanView software and Beaufort Lagoons Ecosystem LTER field protocols. 

### Load required packages

```{r setup}

library(devtools)
devtools::install_github("bristol-emily/CDOMtools")
library(CDOMtools)

```

### Read in a batch of files

Read wavelength and absorbance data from a folder of .txt files. This example uses .txt files bundled in CDOMtools and stored in inst/extdata.

```{r read files}

library(readr)
library(dplyr)
library(purrr)

paths <- dir(
  path = system.file("extdata", package = "CDOMtools", mustWork = TRUE), 
  pattern = "\\.txt$", 
  full.names = TRUE)
names(paths) <- basename(paths)
raw_cdom <- map_df(paths, 
                   read_table, 
                   skip = 14, # skip the header of the text file
                   col_names = FALSE, # .txt does not include column names
                   .id = "filename")

```



```{r}
## Assign clear column names          
raw_cdom <- raw_cdom %>%
  dplyr::rename(wavelength = "X1", # rename wavelength column 
         absorbance = "X2") # rename absorbance column

```

### Interpolate raw absorbance data to integer wavelengths

Use linear interpolation to calculate absorbance at integer wavelengths

```{r interpolation}

## Iterate interpolation function over each scan
interpolated_cdom <- raw_cdom %>% 
  split(.$filename) %>% # divide the vector by file
  map_dfr( # iterate interpolation function over each file
    ~interpolate_absorbance(.$wavelength, # wavelength vector
                            .$absorbance), # absorbance vector
    .id="filename") # keep filename as id

```

### Include timestamp data

To look for instrument drift or zeroing problems during the QAQC process, it is useful to extract the timestamp from the .txt file. Note that this timestamp should only be using for correcting scans and not as the date of sample collection, which is typically included in the file name.

```{r timestamp}

## Extract date and time data from raw .txt files
timestamp <- map_df(paths,
                    read_table, 
                    skip = 2, 
                    n_max = 1, 
                    col_names = FALSE, 
                    .id = "filename", 
                    cols(.default = "c")) 

## Join data to create a column that includes the date and time
timestamp <- timestamp %>% 
  unite(X7, X3, X4, X5, col = datetime, sep = " ") %>% # join and rename year, month, day, time
  rename(tz = X6) %>% # rename column with timezone data
  select(-c(X1, X2))  # remove unnecessary columns

## Check if timezone is same for all data
length(unique(timestamp$tz))

## If there are multiple timezones in the dataset, convert to datetime type using case_when
## See OlsonNames() and timezones() for info on how to set the timezone
timestamp <- timestamp %>% 
  mutate(datetime = case_when(
    tz == "CDT" ~ as_datetime(datetime, "America/Chicago"),
    tz == "AKDT" ~ as_datetime(datetime, "America/Anchorage") 
  ))

## Remove tz column; datetime type data accounts for timezones
timestamp <- select(timestamp, -tz) 

## Add datetime data to interpolated CDOM data
interpolated_cdom <- left_join(interpolated_cdom, timestamp, by = "filename") 

```


### Average triplicate scans

The BLE LTER protocol is to save scans in triplicate. Average the output of the three scans of each sample.

```{r averaging}

## First, create a sample id by removing info about the scan triplicate number (e.g. "..._Absorbance_3")
interpolated_cdom <- interpolated_cdom %>% 
  mutate(sample = sub("_Absorb.*", "", filename)) # remove "Absorbance_x.txt"

## Average the absorbance of triplicate scans and the datetimes of those scans
averaged_cdom <- interpolated_cdom %>% 
  group_by(sample, wavelength) %>% # group by sample and wavelength (n=3)
  summarise(absorbance = mean(absorbance), # average absorbance for each wavelength
            datetime = mean(datetime)) # average datetime from timestamp

View(nest(averaged_cdom, data = c(wavelength, absorbance))) # view averaged samples

```

At this point, data is ready for QA/QC process. Any blanks should be analyzed by day/run to look for instrument drift or issues during the instrument zeroing process. Blanks may be used for correcting sample data. Remove any suspicious looking sample or blank scans.

TODO: Make a vignette about correcting sample data using blank scans.

## Analyzing processed data

### Calculate Napierian absorption cofficient

Absorbance values (i.e. optical density) should be converted to Napierian absorption coefficients for most analyses.

```{r absorption coef}



```

### Calculating spectral slope

### Calculating SUVA data

